---
output: 
  pdf_document:
    latex_engine: xelatex
header-includes:
- \usepackage[fontsize=12pt]{scrextend}
editor_options: 
  markdown: 
    wrap: 72
mainfont: "Times New Roman"
---
\setcounter{page}{14}

# *7.0 Appendix*

```{r setup, results = 'hide', warning=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
heart = read_csv("/Users/yunkyaw/Desktop/School/ADA/heart.csv")

library(tidyr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(corrplot)
library(caret) # confusion matrix, random forest grid search
library(car) # VIF
library(doParallel) # parallel computing
library(glmnet) # logistic regression
library(MASS) # stepwise selection
library(ResourceSelection) # hoslem.test
library(pROC)
library(randomForest)
library(neuralnet)
```


# *7.1 EDA*

## Data Cleaning
```{r, results = 'hide', message = FALSE}
str(heart)

# sex, cp, fbs, restecg, exang, slope, thal, target should all be categorical
heart$sex = as.factor(heart$sex)
heart$cp = as.factor(heart$cp)
heart$fbs = as.factor(heart$fbs)
heart$restecg = as.factor(heart$restecg)
heart$exang = as.factor(heart$exang)
heart$slope = as.factor(heart$slope)
heart$thal = as.factor(heart$thal)
heart$target = as.factor(heart$target)

str(heart)

# factor columns
factor_cols = sapply(heart, is.factor)

# numeric columns
num_cols = sapply(heart, is.numeric)

# determining if there are any missing values
colSums(is.na(heart))
```

## Looking at Distribution of Numeric Variables
```{r, message=FALSE, warning=FALSE}
# convert numeric variables into long data format
melt.heart = melt(heart[num_cols])

# plotting boxplot of each numeric variables to identify outliers and
# distribution
ggplot(data = melt.heart, aes(x = variable, y = value)) + 
  geom_boxplot() +
  labs(title = "Boxplots of Numeric Variables", x = "Variable", y = "Value")
```

## Looking at Distribution of Categorical Variables
```{r, warning = FALSE}
# convert categorical variables into long data format
heart_fact = heart[factor_cols] %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Category")

# plotting proportion of categorical responses
ggplot(heart_fact, aes(x = Category, fill = Category)) +
  geom_bar(aes(y = after_stat(prop), group = 1)) +
  facet_wrap(~ Variable, scales = "free") +
  labs(title = "Distribution of Categorical Variables", y = "Proportion") +
  theme(legend.position = "none")

```

## Correlation Analysis
```{r}
# generate correlation matrix
res = cor(heart[num_cols])

# visualize correlation matrix
corrplot(res, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

## VIF - Assessing Multicollinearity
```{r}
# logistic regression model
log_model = glm(target ~ ., family = "binomial", data = heart)

# evaluate VIF scores
vif_values = vif(log_model)
vif_values
```

# *7.2 Data Analysis*
## Test-Train Partition

```{r}
# making the training set 70% of the data
smp_size = floor(0.7*nrow(heart))

set.seed(123)
train_ind = sample(seq_len(nrow(heart)), size = smp_size)
train = heart[train_ind, ]   # defining training set
test = heart[-train_ind, ]  # defining testing set
```

## Logistic Regression
```{r}
registerDoParallel(cores = 8) # parallel computing 

# find best lambda to apply a penalty to the features
set.seed(123)
cvfit = cv.glmnet(as.matrix(train[, -14]), as.matrix(train[, 14]), nfolds = 10,
                  alpha = 1, family = "binomial", parallel = TRUE)

# final model - logistic regression with lasso penalty
# selecting the lambda that produces the minimum MSE
log_fit = glmnet(as.matrix(train[, -14]), as.matrix(train[, 14]), alpha = 1,
                 family = "binomial", lambda = cvfit$lambda.min)

stopImplicitCluster() # stop parallel computing

# coefficients
coef(log_fit)

# exp(coef) for interpretation
exp(coef(log_fit))
```

### Goodness of Fit
```{r}
# producing probability of each row belonging to the positive class
log_prob = predict(log_fit, newx = as.matrix(test[, -14]))

hoslem.test(as.numeric(test$target), log_prob, g = 10)
# observe good fit as p < 0.05
```

### Generating Predictions and Confusion Matrix
```{r}
# predicting the class of each row with a 0.50 threshold
log_pred = ifelse(log_prob > 0.5, 1, 0)

# confusion matrix
confusionMatrix(as.factor(log_pred), test$target)
```
\newpage

### ROC Curve
```{r, warning = FALSE, message = FALSE}
# developing and plotting ROC curve
roc_obj = roc(response = test$target, predictor = log_prob, type = "response")

plot(roc_obj, legacy.axes = FALSE, print.auc = TRUE, print.three = TRUE,
     cex.lab = 2)
```

```{r}
# calculating area under the curve
AUC = auc(roc_obj)

# getting threshold, sensitivity, accuracy, and AUC for current model
roc_log = c(coords(roc_obj, "b", ret = c("threshold", "se", "sp", "accuracy"), 
                   best.method = "youden"), AUC)

names(roc_log) = c("Threshold", "Sensitivity", "Specificity", "Accuracy", "AUC")

t(roc_log)
```

### Changing Threshold
```{r}
log_pred_adj = ifelse(log_prob > 0.3762408, 1, 0)

# accuracy
log_accuracy = mean(log_pred_adj == test[, 14])

# confusion matrix
confusionMatrix(as.factor(log_pred_adj), test$target)
```
\newpage

## Random Forest
```{r}
registerDoParallel(cores = 8) # parallel computing 

set.seed(123)

# control function for grid search, setting 10-fold cv repeated 3 times
control = trainControl(method = "repeatedcv", number = 10, repeats = 3,
                       search = "grid")

# defining the grid parameters - mtry (number of splits per tree)
grid = expand.grid(mtry = c(1:10))

# training the random forest model with 10-fold cross validation and 1000 trees
rf_gridsearch = train(target~., data=train, method="rf", tuneGrid = grid,
                      trControl = control)

plot(rf_gridsearch)
# determine optimal mtry = 9

# optimal random forest 
rf_gridsearch
```

```{r}
# Generating Optimal Model
rfModel = randomForest(target ~ ., data = train, proximity = T, ntree = 1000, mtry = 9) 
stopImplicitCluster() # stop parallel computing
```

```{r}
rfModel
```

### Visualize Important Features
```{r}
# ranked features by importance using mean decrease gini index
varImpPlot(rfModel)
```

### Generating Predictions and Confusion Matrix
```{r}
rf_prob = predict(rfModel, newdata = test[, -14], type = "prob")

rf_pred = predict(rfModel, newdata = test[, -14], type = "response")

confusionMatrix(rf_pred, test$target)
```

## Neural Network
```{r}
# train: scale numeric columns
train[num_cols] = lapply(train[num_cols], function(x) scale(x))

# train: convert all factor columns to numeric
train[factor_cols] = lapply(train[factor_cols], function(x)
  as.numeric(as.character(x)))

# test: scale numeric columns
test[num_cols] = lapply(test[num_cols], function(x) scale(x))

# test: convert all factor columns to numeric
test[factor_cols] = lapply(test[factor_cols], function(x) as.numeric(as.character(x)))
```

### Grid Search to Find Optimal Layers
```{r}
# generate neural network using grid search to find the optimal number of
# layers and nodes

# generating grid of hidden layer combinations
grid = list(c(5), c(10), c(5, 3), c(10, 5), c(10, 10))

# to store accuracies of each combination
results = data.frame()

registerDoParallel(cores = 8) # parallel computing 
for (i in grid) {
  set.seed(123)
  nn_model = neuralnet(target ~ ., data = train, hidden = i,linear.output = FALSE)
  
  # predict
  nn_prob = predict(nn_model, newdata = test[, -14], type = "prob")
  
  nn_pred = ifelse(nn_prob > 0.5, 1, 0)
  
  # accuracy
  acc = mean(nn_pred == test[, 14])
  
  results = rbind(results, data.frame(Hidden = paste(i, collapse = "-"),
                                      Accuracy = acc))
}
stopImplicitCluster() # stop parallel computing
```

```{r}
results
# find best results with one layer and 10 neurons
```

### Generating Optimal Neural Network
```{r}
nn_model = neuralnet(target ~ ., data = train, hidden = 10, linear.output = FALSE)
```

### Visualize Optimal Neural Network
```{r}
plot(nn_model, rep="best")
```

###  Generating Predictions and Confusion Matrix
```{r}
nn_prob = predict(nn_model, newdata = test[, -14], type = "prob")

nn_pred = ifelse(nn_prob > 0.5, 1, 0)

# confusion matrix
confusionMatrix(as.factor(nn_pred), as.factor(test$target))
```

# *7.3 Comparison*
## ROC Curves
```{r, warning = FALSE, message=FALSE}
# logistic regression
roc_log = roc(response = test$target, predictor = log_prob)
# AUC = 0.9254

# random forest - looking at the probability of a positive observation
roc_rf = roc(response = test$target, predictor = rf_prob[, "1"])
# AUC = 1

# neural network
roc_nn = roc(response = test$target, predictor = nn_prob)
# AUC = 0.9834

# plotting
plot(roc_log, legacy.axes = FALSE, cex.lab = 2, col = "green")
lines(roc_log, col = "green")
lines(roc_rf, col = "red")
lines(roc_nn, col = "blue")
legend("bottomright", legend = c("Logistic Regression (AUC = 0.9254)",
                                 "Random Forest (AUC = 1)",
                                 "Neural Network (AUC = 0.9834)"), 
       col = c("green", "red", "blue"), lty = 1)
```
\newpage

## Table of Accuracies
```{r}
# retrieved metric values from confusionMatrix for each model
metrics = c("Accuracy", "Sensitivity", "Specificity", "AUC")
log_metr = c(0.8766, 0.8759, 0.8773, 0.9254)
rf_metr = c(0.9903, 1, 0.9816, 1)
nn_metr = c(0.9838, 0.9793, 0.9877, 0.9834)

results_table <- data.frame(
  Metric = metrics,
  Logistic_Regression = log_metr,
  Random_Forest = rf_metr,
  Neural_Network = nn_metr
)

t(results_table)
```

